This page and the next page should be removed

### 2021-01-14 analysis/manuscript.Rmd
implement feedback by rasha and garrett

author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
  - name          : "Franziska Süß"
    affiliation   : "3"
  - name          : "Rasha Abdel Rahman"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Rudower Chaussee 18, 12489 Berlin"
    email         : "rasha.abdel.rahman@hu-berlin.de"

affiliation:
  - id            : "1"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "2"
    institution   : "Max Planck Institute for Human Cognitive and Brain Sciences, Leipzig, Germany"
  - id            : "3"
    institution   : "Fachhochschule des Mittelstands, Bamberg, Germany"

@@

keywords          : "objects, semantic knowledge, visual perception, event-related potentials"
wordcount         : ""

bibliography      : "manuscript_files/r-references.bib"
zotero            : "aha"

@@

csl               : "manuscript_files/apa.csl"
documentclass     : "apa7"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine  : "xelatex"

header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
---

@@

How does our perception of an object change as soon as we discover what it is for? In this study, we presented participants with unfamiliar objects which were preceded either by valid information about their function, leading to semantically informed perception, or by misleading information, leading to naive perception. Semantically informed perception of objects was expected to differ from naive perception at certain stages in the processing hierarchy and we aimed to identify these stages by capitalizing on the high temporal resolution of event-related brain potentials (ERPs).

@@

Empirical findings from experimental psychology have recently added additional support to the view that perception can indeed be penetrated by other aspects of cognition [for review, see @collins2014; @vetter2014]. These aspects of cognition may include transient states such as emotions [e.g., @bocanegra2009; @phelps2016] or intentions [e.g., @balcetis2010; @cole2012] as well as learned capacities such as language [e.g., @boutonnet2015; @maier2018; @mo2011] or declarative knowledge about faces [e.g., @abdelrahman2011; @suess2015; @eiserbeck2020] and objects [e.g., @abdelrahman2008; @lin1997; @weller2019]. Despite this wealth of empirical evidence, critics of the idea that cognitive functions can penetrate perception have remained skeptical, pointing out important methodological shortcomings in large parts of this literature [@firestone2016; @machery2015].

One productive way to investigate the top-down effects of cognition on perception while addressing some of these concerns is to use neurophysiological recording methods such as ERPs [@luck2014]: Their high temporal resolution allows for a direct test of which processing stages---perceptual and/or postperceptual---are influenced by higher-level capacities such as semantic knowledge. This is made possible by the fact that when participants are presented with visual stimuli, their averaged brain responses show characteristic deflections (ERP components), each of them with its typical polarity (positive or negative), peak latency, spatial scalp distribution, and functional role(s).

@@

In the two experiments subsequently reported, participants were presented with real-word objects that were presumed to be unfamiliar to most of them. They first viewed each of these objects without any semantic information. This first part served as a naive baseline to rule out that ERPs would differ in response to the objects based solely on low-level visual differences. Next, participants viewed each unfamiliar object for a second time, now preceded by verbal keywords. These keywords could either correctly describe the typical function of the object, thus making it possible for participants to understand what kind of object they were viewing, or they could mislead participants by describing the function of a different object, thus keeping the perception of the object semantically naive. During this second part, we were able measure the online influence of semantic insight on object-evoked ERPs as it happened. Finally, the objects were presented for a third time (without keywords, as in the first part) to investigate downstream effects of having acquired semantic knowledge about them---thereby mimicking previous studies [e.g., @abdelrahman2008; @samaha2018; @weller2019]. In all three parts, we examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component, 100--150 ms), higher-level visual perception (N170 component, 150--200 ms), and semantic processing (N400 component, 400--700 ms).

@@

Single-trial ERPs were analyzed in response to unfamiliar objects before (pre-insight part), while (insight part), and after (post-insight part) participants obtained relevant semantic information about their function. Only in the insight part, half of the objects were preceded by matching keywords, fostering semantically informed perception to occur, whereas the other half were preceded by non-matching keywords, keeping the perception of the object semantically naive. The objects were analyzed according to this manipulation in combination with participants' self-report in the insight part (see Figure \@ref(fig:exp1-plot)A), thereby making sure that semantically informed and naive perception did indeed occur. The analysis focused on differences between these two conditions in the P1 component (100--150 ms) as an index of lower-level visual perception, the N170 component (150--200 ms) as index of higher-level visual perception, and the N400 component (400--700 ms) as an index of semantic processing.

@@

Participants for Experiment 2 were 24 German native speakers (15 female, 9 male) with a mean age of 26 years (range 19 to 29 years) who had not participated in Experiment 1. They had no history of psychological disorder or treatment, were right-handed, and reported normal or corrected-to-normal vision. They gave written informed consent before starting the experiment and received a compensation of €8 per hour for participating.

@@

One may raise concerns whether the modulation of the N170 component in the insight part genuinely reflects the semantically informed perception of the objects in the respective condition, or---as an alternative explanation---whether it may be driven by the objects in the other, semantically naive condition. Remember that these objects were preceded by non-matching keywords which were picked so that they could not be related to the visual features of the object and their configuration. Thus, the modulation of the ERP components in the insight part could be a mismatch response to those objects---reflecting, for example, the fact that the visual features of the object shown in Figure \@ref(fig:exp1-plot)A cannot be reconciled with the function of morsing messages. To preclude this alternative explanation, we repeated our analysis using a different baseline against which the objects in the semantically informed condition were contrasted. Instead of the naive condition (where objects were presented with non-matching keywords), we now used those objects which were preceded by matching keywords (as in the informed condition), but which were excluded from the main analysis because participants indicated behaviorally that they did not understand the object they were seeing. Across both experiments, this was the case for `r scales::percent(stims_joint["Excl_informed"]/60, accuracy = 0.1)` of objects presented with matching keywords as compared to `r scales::percent(stims_joint["Informed"]/60, accuracy = 0.1)` of objects which did indeed lead to semantically informed perception.[^1] Just as above, this control analysis revealed a robust N170 effect in the insight part, *b* = `r table_joint_control$conts["Insight part", "N170_est"]` µV, *p* `r table_joint_control$conts["Insight part", "N170_p"]`. Thus, this enhanced negativity seems to be a genuine marker of semantically informed perception, no matter if compared to objects presented with misleading keywords or compared to objects presented with accurate keywords but on which participants failed to capitalize. Note that the reduction of the N400 component in the insight part also remained robust in this control analysis, *b* = `r table_joint_control$conts["Insight part", "N400_est"]` µV, *p* `r table_joint_control$conts["Insight part", "N400_p"]`.

@@

Here we investigated if obtaining a semantic understanding of previously unfamiliar objects has an influence on how we perceive them. To this end, we measured ERPs while participants viewed unfamiliar objects before, while, and after receiving semantic information about them. For half of the objects, this information was matching the object, thus leading to semantically informed perception, whereas for the other half of the objects, it was misleading, thus keeping the perception of the object semantically naive. We found semantically informed perception to be accompanied by enlarged (i.e., more negative) N170 amplitudes and reduced (i.e., less negative) N400 amplitudes. When the same objects were presented again without the semantic information, the N400 component remained significantly reduced and we also observed a modulation of the P1 component, which was enlarged (i.e., more positive) in response to objects that had previously triggered semantically informed perception.

>>>>

author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
  - name          : "Franziska Süß"
    affiliation   : "3"
  - name          : "Rasha Abdel Rahman"
    affiliation   : "1"
    corresponding : yes
    address       : "Rudower Chaussee 18, 12489 Berlin"
    email         : "rasha.abdel.rahman@hu-berlin.de"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Humboldt-Universität zu Berlin"
  - id            : "2"
    institution   : "Research Group Learning in Early Childhood, Max Planck Institute for Human Cognitive and Brain Sciences, Leipzig, Germany"
  - id            : "3"
    institution   : "Fachhochschule des Mittelstands, Bamberg, Germany"

@@

keywords          : "objects, semantic knowledge, visual perception, event-related potentials"

@@

bibliography      : "manuscript_files/r-references.bib"
csl               : "manuscript_files/apa.csl"
zotero            : "aha"
documentclass     : "apa7"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine  : "xelatex"

header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
---

@@

How does our perception of an object change as soon as we discover what it is for? In this study, we presented participants with unfamiliar objects which were preceded either by matching information about their function, leading to semantically informed perception, or by non-matching information, leading to naive perception. Semantically informed perception of objects was expected to differ from naive perception at certain stages in the processing hierarchy and we aimed to identify these stages by capitalizing on the high temporal resolution of event-related brain potentials (ERPs).

@@

Empirical findings from experimental psychology have recently added additional support to the view that perception can indeed be penetrated by other aspects of cognition [for review, see @collins2014; @vetter2014]. These aspects of cognition may include transient states such as emotions [e.g., @bocanegra2009; @phelps2016] or intentions [e.g., @balcetis2010; @cole2012] as well as learned capacities such as language [e.g., @boutonnet2015; @maier2018; @mo2011] or declarative knowledge about faces [e.g., @abdelrahman2011; @suess2015; @eiserbeck2020] and objects [e.g., @abdelrahman2008; @lin1997; @weller2019]. Despite this wealth of empirical evidence, critics of the idea that cognitive functions can penetrate perception have remained skeptical and pointed out important methodological shortcomings in large parts of this literature [@firestone2016; @machery2015]. They have argued, for example, that most behavioral paradigms have not been able to discern between perceptual and postperceptual (e.g., memory-related) effects. Furthermore, cognitive influences in some studies have been confounded with differences in the lower-level visual input. We will revisit these and other potential pitfalls with regards to the present study in the Discussion section once we have laid out our methodology and results.

In general, however, it is important to note that one productive way to circumvent some of these concerns is to use neurophysiological recording methods such as ERPs [@luck2014]: Their high temporal resolution allows for a direct test of which processing stages---perceptual and/or postperceptual---are influenced by higher-level capacities such as semantic knowledge. This is made possible by the fact that when participants are presented with visual stimuli, their averaged brain responses show characteristic deflections (ERP components), each of them with its typical polarity (positive or negative), peak latency, spatial scalp distribution, and functional role(s).

@@

In the two experiments subsequently reported, participants were presented with real-word objects that were presumed to be unfamiliar to most of them. They first viewed each of these objects without any semantic information. This first part served as a naive baseline to rule out that ERPs would differ in response to the objects based solely on low-level visual differences. Next, participants viewed each unfamiliar object for a second time, now preceded by verbal keywords. These keywords could either be matching the typical function of the object, thus making it possible for participants to understand what kind of object they were viewing, or they could be non-matching by describing the function of a different object, thus keeping the perception of the object semantically naive. During this second part, we were able measure the online influence of semantic insight on object-evoked ERPs as it happened. Finally, the objects were presented for a third time (without keywords, as in the first part) to investigate downstream effects of having acquired semantic knowledge about them---thereby mimicking previous studies [e.g., @abdelrahman2008; @samaha2018; @weller2019]. In all three parts, we examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component, 100--150 ms), higher-level visual perception (N170 component, 150--200 ms), and semantic processing (N400 component, 400--700 ms).

@@

Single-trial ERPs were analyzed in response to unfamiliar objects before (pre-insight part), while (insight part), and after (post-insight part) participants obtained relevant semantic information about their function. In the insight part, half of the objects were preceded by matching keywords, fostering semantically informed perception. The other half were preceded by non-matching keywords, keeping the perception of the object semantically naive. The objects were analyzed according to this manipulation in combination with participants' self-report in the insight part (see Figure \@ref(fig:exp1-plot)A), thereby making sure that semantically informed and naive perception did indeed occur. The analysis focused on differences between these two conditions in the P1 component (100--150 ms) as an index of lower-level visual perception, the N170 component (150--200 ms) as index of higher-level visual perception, and the N400 component (400--700 ms) as an index of semantic processing.

@@

Participants for Experiment 2 were 24 German native speakers (15 female, 9 male) with a mean age of 26 years (range 19 to 29 years) who had not participated in Experiment 1. They had no history of psychological disorders or treatment, were right-handed, and reported normal or corrected-to-normal vision. They gave written informed consent before starting the experiment and received a compensation of €8 per hour for participating.

@@

One may raise concerns whether the modulation of the N170 component in the insight part genuinely reflects the semantically informed perception of the objects in the respective condition, or---as an alternative explanation---whether it may be driven by the objects in the other, semantically naive condition. Remember that these objects were preceded by non-matching keywords which were picked so that they could not be related to the visual features of the object and their configuration. Thus, the modulation of the ERP components in the insight part could be a mismatch response to those objects---reflecting, for example, the fact that the visual features of the object shown in Figure \@ref(fig:exp1-plot)A cannot be reconciled with the function of morsing messages. To preclude this alternative explanation, we repeated our analysis using a different baseline against which the objects in the semantically informed condition were contrasted. Instead of the naive condition (where objects were presented with non-matching keywords), we now used those objects which were preceded by matching keywords (as in the informed condition), but which were excluded from the main analysis because participants indicated behaviorally that they did not understand the object they were seeing. Across both experiments, this was the case for `r scales::percent(stims_joint["Excl_informed"]/60, accuracy = 0.1)` of objects presented with matching keywords as compared to `r scales::percent(stims_joint["Informed"]/60, accuracy = 0.1)` of objects which did indeed lead to semantically informed perception.[^1] Just as above, this control analysis revealed a robust N170 effect in the insight part, *b* = `r table_joint_control$conts["Insight part", "N170_est"]` µV, *p* `r table_joint_control$conts["Insight part", "N170_p"]`. Thus, this enhanced negativity seems to be a genuine marker of semantically informed perception, no matter if compared to objects presented with non-matching keywords or compared to objects presented with matching keywords but on which participants failed to capitalize. Note that the reduction of the N400 component in the insight part also remained robust in this control analysis, *b* = `r table_joint_control$conts["Insight part", "N400_est"]` µV, *p* `r table_joint_control$conts["Insight part", "N400_p"]`.

@@

Here we investigated if obtaining a semantic understanding of previously unfamiliar objects has an influence on how we perceive them. To this end, we measured ERPs while participants viewed unfamiliar objects before, while, and after receiving semantic information about them. For half of the objects, this information was matching the object, thus leading to semantically informed perception, whereas for the other half of the objects, it was non-matching, thus keeping the perception of the object semantically naive. We found semantically informed perception to be accompanied by enlarged (i.e., more negative) N170 amplitudes and reduced (i.e., less negative) N400 amplitudes. When the same objects were presented again without the semantic information, the N400 component remained significantly reduced and we also observed a modulation of the P1 component, which was enlarged (i.e., more positive) in response to objects that had previously triggered semantically informed perception.
